---
title: "Corrected Standard Errors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{corrected-se}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
header-includes:
  - \DeclareMathOperator{\Var}{\mathrm{Var}}
  - \DeclareMathOperator{\E}{\mathrm{E}}
  - \newcommand{\bv}[1]{\boldsymbol{\mathbf{#1}}}  %APA-consistent bold
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how to obtain Taylor-series corrected standard errors in 2S-PA, accounting for the uncertainty in the weights for obtaining the factor scores. It is described in [this paper](https://doi.org/10.1007/s00181-020-01942-z).

The second-stage estimation treats as known the loading matrix and the error covariance matrix of the factor scores as indicators of the true latent latent variables. However, those are estimates based on the first-stage measurement model. When the sample size is large and/or the reliability of the factor scores are high, the uncertainty in the estimated loading and error covariance matrix is generally negligible. Otherwise, a first-order correction on the standard errors of the second-stage estimation is possible, as illustrated below.

First-order correction for SE

$$\hat V_{\gamma, c} = \hat V_{\gamma} + (\boldsymbol{H}_\gamma)^{-1} \left(\frac{\partial^2 \ell}{\partial \theta \partial \gamma^\top}\right) \hat V_{\theta} \left(\frac{\partial^2 \ell}{\partial \theta \partial \gamma^\top}\right)^\top (\boldsymbol{H}_\gamma)^{-1},$$

where $V_{\gamma}$ is the naive covariance matrix of the structural parameter estimates $\hat{\boldsymbol{\gamma}}$ assuming the measurement error variance parameter, $\boldsymbol{\theta}$, is known, $\boldsymbol{H}_\gamma$ is the Hessian matrix of the log-likelihood $\ell$ with respect to $\hat{\boldsymbol{\gamma}}$, and $V_{\theta}$ can be obtained in the first-stage measurement model analysis.

```{r load-pkg, message = FALSE}
library(lavaan)
library(R2spa)
library(numDeriv)
library(boot)
```

This example is based on the Political Democracy data used in <https://lavaan.ugent.be/tutorial/sem.html> and as an example in Bollen (1989)'s book.

## Separate Measurement Models

```{r}
model <- '
  # latent variable definitions
     ind60 =~ x1 + x2 + x3
     dem60 =~ y1 + y2 + y3 + y4

  # regressions
    dem60 ~ ind60
'
```

```{r}
cfa_ind60 <- cfa("ind60 =~ x1 + x2 + x3", data = PoliticalDemocracy)
# Regression factor scores
fs1 <- get_fs_lavaan(cfa_ind60, vfsLT = TRUE)
# Delta method variance of (loading, intercept)
vldev1 <- attr(fs1, which = "vfsLT")
cfa_dem60 <- cfa("dem60 =~ y1 + y2 + y3 + y4",
                 data = PoliticalDemocracy)
# Regression factor scores
fs2 <- get_fs_lavaan(cfa_dem60, vfsLT = TRUE)
# Delta method variance of (loading, intercept)
vldev2 <- attr(fs2, which = "vfsLT")
fs_dat <- data.frame(
  fs_ind60 = fs1$fs_ind60,
  fs_dem60 = fs2$fs_dem60
)
# Combine sampling variance of loading and error variance
# Note: loadings first, then error variance
vldev <- block_diag(vldev1, vldev2)[c(1, 3, 2, 4), c(1, 3, 2, 4)]
```

```{r}
# 2S-PA
# Assemble loadings
ld <- block_diag(attr(fs1, which = "fsL"),
                 attr(fs2, which = "fsL"))
ev <- block_diag(attr(fs1, which = "fsT"),
                 attr(fs2, which = "fsT"))
tspa_fit <- tspa(model = "dem60 ~ ind60",
                 data = fs_dat,
                 cross_loadings = ld,
                 vc = ev)
# Unadjusted covariance
vcov(tspa_fit)
# Adjusted covariance matrix
(vc_cor <- vcov_corrected(tspa_fit, vfsLT = vldev, which_free = c(1, 4, 5, 7)))
# Corrected standard errors
sqrt(diag(vc_cor))
```

Standardized solution

```{r, eval = FALSE}
tspa_est_std <- function(se) {
  tfit <- tspa(model = "dem60 ~ ind60",
               data = fs_dat,
               se = list(ind60 = se[1], dem60 = se[2]))
  standardizedSolution(tfit)$est.std[5]
}
tspa_est_std(unlist(se_fs))
# Jacobian
jac_std <- jacobian(tspa_est_std, x = unlist(se_fs))
# Corrected standard error
sqrt(
  standardizedSolution(tspa_fit)[5, "se"]^2 +
    jac_std %*% vc_ev_fs %*% t(jac_std)
)
```

Compare to joint model

```{r, eval = FALSE}
sem_fit <- sem(model, data = PoliticalDemocracy)
# Larger standard error
vcov(sem_fit)[c("dem60~ind60", "ind60~~ind60", "dem60~~dem60"),
              c("dem60~ind60", "ind60~~ind60", "dem60~~dem60")]
```

Bootstrap Standard Errors

```{r, eval = FALSE, echo = TRUE}
run_tspa <- function(df, inds) {
  fs_ind60 <- get_fs(data = df[inds, ], model = "ind60 =~ x1 + x2 + x3",
                     se = "none", test = "none")
  fs_dem60 <- get_fs(data = df[inds, ], model = "dem60 =~ y1 + y2 + y3 + y4",
                     se = "none", test = "none")
  fs_dat <- cbind(fs_ind60, fs_dem60)
  # Assemble loadings
  ld <- block_diag(attr(fs_ind60, which = "fsL"),
                   attr(fs_dem60, which = "fsL"))
  ev <- block_diag(attr(fs_ind60, which = "fsT"),
                   attr(fs_dem60, which = "fsT"))
  tspa_fit <- tspa(model = "dem60 ~ ind60",
                   data = fs_dat,
                   cross_loadings = ld,
                   vc = ev,
                   test = "none")
  coef(tspa_fit)
}
boo <- boot(PoliticalDemocracy, statistic = run_tspa, R = 1999)
```

```{r, eval = FALSE, include = FALSE}
saveRDS(boo, file = "boo_separate.RDS")
```

```{r, include = FALSE}
boo <- readRDS("boo_separate.RDS")
```

```{r}
# Use MAD to downweigh outlying replications
boo$t |>
    apply(MARGIN = 2, FUN = mad) |>
    setNames(c("dem60~ind60", "ind60~~ind60", "dem60~~dem60"))
```

```{r, eval = FALSE}
run_sem <- function(df, inds) {
  sem_fit <- sem(model, data = df[inds, ], se = "none", test = "none")
  coef(sem_fit)
}
boo <- boot(PoliticalDemocracy, statistic = run_sem, R = 999)
```

## Joint Measurement Model

```{r}
cfa_joint <- cfa("ind60 =~ x1 + x2 + x3
                  dem60 =~ y1 + y2 + y3 + y4",
                 data = PoliticalDemocracy)
# Factor score
fs_joint <- get_fs_lavaan(cfa_joint, vfsLT = TRUE)
# Delta method variance
vldev_joint <- attr(fs_joint, which = "vfsLT")
```

```{r}
tspa_fit2 <- tspa(model = "dem60 ~ ind60",
                  data = fs_joint,
                  vc = attr(fs_joint, "fsT"),
                  cross_loadings = attr(fs_joint, "fsL"))
# Unadjusted covariance
vcov(tspa_fit2)
# Adjusted covariance
(vc2_cor <- vcov_corrected(tspa_fit2, vfsLT = vldev_joint))
# Corrected standard errors
sqrt(diag(vc2_cor))
```

Bootstrap Standard Errors

```{r, eval = FALSE}
run_tspa2 <- function(df, inds) {
  fs_joint <- get_fs(df[inds, ],
                     model = "ind60 =~ x1 + x2 + x3
                              dem60 =~ y1 + y2 + y3 + y4",
                     se = "none", test = "none")
  tspa_fit2 <- tspa(model = "dem60 ~ ind60",
                    data = fs_joint,
                    vc = attr(fs_joint, "fsT"),
                    cross_loadings = attr(fs_joint, "fsL"),
                    test = "none")
  coef(tspa_fit2)
}
boo2 <- boot(PoliticalDemocracy, statistic = run_tspa2, R = 1999)
# run_tspa2b <- function(df, inds) {
#   fsb_joint <- get_fs(df[inds, ],
#                       model = "ind60 =~ x1 + x2 + x3
#                               dem60 =~ y1 + y2 + y3 + y4",
#                       se = "none", test = "none", method = "Bartlett")
#   tspa_fit2b <- tspa(model = "dem60 ~ ind60",
#                      data = fsb_joint,
#                      vc = attr(fsb_joint, "fsT"),
#                      cross_loadings = diag(2),
#                      test = "none")
#   coef(tspa_fit2b)
# }
# boo2b <- boot(PoliticalDemocracy, statistic = run_tspa2b, R = 4999)
# run_sem2 <- function(df, inds) {
#   sem_fit <- sem(model, data = df[inds, ])
#   coef(sem_fit)[c(6, 14, 15)]
# }
# boo2j <- boot(PoliticalDemocracy, statistic = run_sem2, R = 4999)
```

```{r, eval = FALSE, include = FALSE}
saveRDS(boo2, file = "boo_joint.RDS")
```

```{r, include = FALSE}
boo2 <- readRDS("boo_joint.RDS")
```

```{r}
# Use MAD to downweigh outlying replications
boo2$t |>
    apply(MARGIN = 2, FUN = mad) |>
    setNames(c("dem60~ind60", "ind60~~ind60", "dem60~~dem60"))
```

### With Bartlett's Method

```{r}
# Factor score
fsb_joint <- get_fs(PoliticalDemocracy,
                    model = "ind60 =~ x1 + x2 + x3
                             dem60 =~ y1 + y2 + y3 + y4",
                    method = "Bartlett",
                    vfsLT = TRUE)
vldevb_joint <- attr(fsb_joint, which = "vfsLT")
```

```{r}
tspa_fit2b <- tspa(model = "dem60 ~ ind60",
                   data = fsb_joint,
                   vc = attr(fsb_joint, "fsT"),
                   cross_loadings = diag(2) |>
                       `dimnames<-`(list(c("fs_ind60", "fs_dem60"),
                                         c("ind60", "dem60"))))
# Unadjusted covariance matrix
vcov(tspa_fit2b)
# Adjusted covariance matrix
(vc2b_cor <- vcov_corrected(
    tspa_fit2b,
    # Exclude fixed loadings and error variance
    vfsLT = vldevb_joint[c(5, 7), c(5, 7)],
    # Specify which elements are free (error variances only)
    which_free = c(5, 7)))
# Corrected standard errors
sqrt(diag(vc2b_cor))
```

## Multiple Groups

```{r}
# Multigroup, three-factor example
mod <- "
  # latent variables
    visual =~ x1 + x2 + x3
    textual =~ x4 + x5 + x6
    speed =~ x7 + x8 + x9
"
# Factor scores based on partial invariance
fs_dat <- get_fs(HolzingerSwineford1939, model = mod, std.lv = TRUE,
                 group = "school",
                 group.equal = c("loadings", "intercepts"),
                 group.partial = c("visual=~x2", "x7~1"))
fit <- cfa(mod,
           data = HolzingerSwineford1939,
           std.lv = TRUE,
           group = "school",
           group.equal = c("loadings", "intercepts"),
           group.partial = c("visual=~x2", "x7~1"))
fs_dat <- get_fs_lavaan(fit)
vldev <- R2spa:::vcov_ld_evfs(fit)
```

```{r}
tspa_fit <- tspa(model = "visual ~~ textual + speed
                          textual ~~ speed",
                 data = fs_dat,
                 group = "school",
                 cross_loadings = attr(fs_dat, which = "fsL"),
                 vc = attr(fs_dat, which = "fsT"))
# Unadjusted covariance
vcov(tspa_fit)[c(1:6, 10:15), c(1:6, 10:15)]
# Adjusted covariance
vcov_corrected(tspa_fit, vfsLT = vldev)[c(1:6, 10:15), c(1:6, 10:15)]
```
