---
title: "Error vs. Reliability Constraints"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{error-vs-reliability}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# library(R2spa)
library(lavaan)
```

The example is from https://lavaan.ugent.be/tutorial/sem.html. 

# Full SEM

```{r}
myModel <- '
   # latent variables
     ind60 =~ x1 + x2 + x3
     dem60 =~ y1 + y2 + y3 + y4
   # regressions
     dem60 ~ ind60
'
fit <- sem(model = myModel,
           data  = PoliticalDemocracy)
summary(fit)
semPlot::semPaths(fit)
standardizedSolution(fit)
```

# 2S-PA With Two-Factor CFA

## Stage 1: CFA

```{r}
# CFA
my_cfa <- '
   # latent variables
     ind60 =~ x1 + x2 + x3
     dem60 =~ y1 + y2 + y3 + y4
'
cfa_fit <- cfa(model = my_cfa,
               data  = PoliticalDemocracy,
               std.lv = TRUE)
# Compute factor scores using lavPredict
fs_dat <- lavPredict(cfa_fit, acov = "standard")
colnames(fs_dat) <- c("fs_ind60", "fs_dem60")
# Alpha reliability
alpha_x <- psych::alpha(PoliticalDemocracy[c("x1", "x2", "x3")])$total$raw_alpha
alpha_y <- psych::alpha(PoliticalDemocracy[c("y1", "y2", "y3", "y4")])$total$raw_alpha
```

## Stage 2

### Fixing errors

```{r}
my2spa <- '
   # latent variables (indicated by factor scores)
     ind60 =~ 1 * fs_ind60
     dem60 =~ 1 * fs_dem60
   # constrain the errors
     fs_ind60 ~~ 0.036168769 * fs_ind60
     fs_dem60 ~~ 0.104325893 * fs_dem60
     fs_ind60 ~~ 0.002202184 * fs_dem60
   # latent variances
     ind60 ~~ v1 * ind60
     dem60 ~~ v2 * dem60
   # regressions
     dem60 ~ ind60
'
tspa_fit1 <- sem(model = my2spa, data  = fs_dat)
standardizedSolution(tspa_fit1)
```

### Constraining Reliability

```{r}
# Reliability
cfa_psi <- lavInspect(cfa_fit, what = "est")$psi
1 - attr(fs_dat, "se")[[1]]^2 / diag(cfa_psi)
my2spa <- '
   # latent variables (indicated by factor scores)
     ind60 =~ 1 * fs_ind60
     dem60 =~ 1 * fs_dem60
   # constrain the errors
     fs_ind60 ~~ ev1 * fs_ind60
     fs_dem60 ~~ ev2 * fs_dem60
   # latent variances
     ind60 ~~ v1 * ind60
     dem60 ~~ v2 * dem60
   # regressions
     dem60 ~ b1 * ind60
   # reliability constraints (reliability = v / (v + ev))
   # v = reliability / (1 - reliability) * ev
     v1 == 0.9638312 / (1 - 0.9638312) * ev1
     v2 == 0.895674 / (1 - 0.895674) * ev2 - b1^2 * v1
'
tspa_fit2 <- sem(model = my2spa, data  = fs_dat)
standardizedSolution(tspa_fit2)
```

# 2S-PA With Separate One-Factor CFAs

## Stage 1: CFA

```{r}
# CFA
my_cfa1 <- '
   # latent variables
     ind60 =~ x1 + x2 + x3
'
cfa_fit1 <- cfa(model = my_cfa1,
                data  = PoliticalDemocracy,
                std.lv = TRUE)
my_cfa2 <- '
   # latent variables
     dem60 =~ y1 + y2 + y3 + y4
'
cfa_fit2 <- cfa(model = my_cfa2,
                data  = PoliticalDemocracy,
                std.lv = TRUE)
# Compute factor scores using lavPredict
fs1 <- lavPredict(cfa_fit1, se = "standard", acov = "standard")
fs2 <- lavPredict(cfa_fit2, se = "standard", acov = "standard")
fs_dat2 <- data.frame(fs1, fs2)
colnames(fs_dat2) <- c("fs_ind60", "fs_dem60")
```

## Stage 2

### Fixing errors

```{r}
# Error:
# var(fs1) * 0.03423273
# var(fs2) * 0.1131951
my2spasep <- '
   # latent variables (indicated by factor scores)
     ind60 =~ 1 * fs_ind60
     dem60 =~ 1 * fs_dem60
   # constrain the errors
     fs_ind60 ~~ 0.03350762 * fs_ind60
     fs_dem60 ~~ 0.1017384 * fs_dem60
   # latent variances
     ind60 ~~ v1 * ind60
     dem60 ~~ v2 * dem60
   # regressions
     dem60 ~ ind60
'
tspasep_fit1 <- sem(model = my2spasep, data  = fs_dat2)
standardizedSolution(tspasep_fit1)
```

### Constraining Reliability

```{r}
# Reliability
cfa_psi1 <- lavInspect(cfa_fit1, what = "est")$psi
1 - attr(fs1, "se")[[1]]^2 / diag(cfa_psi1)
cfa_psi2 <- lavInspect(cfa_fit2, what = "est")$psi
1 - attr(fs2, "se")[[1]]^2 / diag(cfa_psi2)
my2spasep <- '
   # latent variables (indicated by factor scores)
     ind60 =~ 1 * fs_ind60
     dem60 =~ 1 * fs_dem60
   # constrain the errors
     fs_ind60 ~~ ev1 * fs_ind60
     fs_dem60 ~~ ev2 * fs_dem60
   # latent variances
     ind60 ~~ v1 * ind60
     dem60 ~~ v2 * dem60
   # regressions
     dem60 ~ b1 * ind60
   # reliability constraints (reliability = v / (v + ev))
   # v = reliability / (1 - reliability) * ev
     v1 == 0.9657673 / (1 - 0.9657673) * ev1
     v2 == 0.8868049 / (1 - 0.8868049) * ev2 - b1^2 * v1
'
tspasep_fit2 <- sem(model = my2spasep, data  = fs_dat2)
standardizedSolution(tspasep_fit2)
```

# Summarize Results

```{r}
extract_b <- function(object) {
  ss <- standardizedSolution(object)
  ss[ss$lhs == "dem60" & ss$op == "~" & ss$rhs == "ind60", 
     c("lhs", "op", "rhs", "est.std", "se", "z", "pvalue",
       "ci.lower", "ci.upper")]
}
coef_bs <- lapply(list(`sem` = fit,
                       `2spa-joint-error` = tspa_fit1,
                       `2spa-joint-reliability` = tspa_fit2,
                       `2spa-sep-error` = tspasep_fit1,
                       `2spa-sep-reliability` = tspasep_fit2),
                  FUN = extract_b)
knitr::kable(do.call(rbind, coef_bs), digits = 3)
```

## Observations

- Inflated latent correlation with the two-factor model for obtaining factor scores. Unclear why this happens. Factor score covariances seem higher than (or equal to) latent covariances, so perhaps attenuation does not apply to factor score covariances?

<!-- ## Reproduce CFA -->

<!-- ```{r} -->
<!-- my2scfa <- ' -->
<!--    # latent variables (indicated by factor scores) -->
<!--      ind60 =~ 1 * fs_ind60 + .05 * fs_dem60 -->
<!--      dem60 =~ 1 * fs_dem60 + .05 * fs_ind60 -->
<!--    # constrain the errors -->
<!--      fs_ind60 ~~ 0.036168769 * fs_ind60 -->
<!--      fs_dem60 ~~ 0.104325893 * fs_dem60 -->
<!--      fs_ind60 ~~ 0.002202184 * fs_dem60 -->
<!--    # latent variances -->
<!--      ind60 ~~ dem60 -->
<!-- ' -->
<!-- tscfa_fit <- cfa(model = my2scfa, data  = fs_dat) -->
<!-- standardizedSolution(tscfa_fit) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- my2scfasep <- ' -->
<!--    # latent variables (indicated by factor scores) -->
<!--      ind60 =~ 1 * fs_ind60 -->
<!--      dem60 =~ 1 * fs_dem60 -->
<!--    # constrain the errors -->
<!--      fs_ind60 ~~ 0.03423273 * fs_ind60 -->
<!--      fs_dem60 ~~ 0.1131951 * fs_dem60 -->
<!--    # latent variances -->
<!--      ind60 ~~ dem60 -->
<!-- ' -->
<!-- tscfasep_fit <- cfa(model = my2scfasep, data  = fs_dat2, -->
<!--                     std.lv = TRUE) -->
<!-- ``` -->
