---
title: "int_ind-example"
author: "Jimmy Zhang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{int_ind-example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r load packages}
library(lavaan)
library(dplyr)
library(R2spa)
```

## `int_ind()` Function
This function is designed to create product indicators for first-order factor score indicators with standard errors and model-implied loadings. 
With a specified model, the product indicator pairs will be generated based on the model syntax; without a specified model, the function will by default generate all possible pairs of product indicators. 

## Illustrative Example

```{r Data Simulation}
# Simulate a dataset with four firsr-order latent variables: x, m, z, y

## Sample size:
num_obs <- 5000

# Structural Parameters: 
## gamma_x = 0.3, gamma_m = 0.4, gamma_z = 0.2
## gamma_xm = 0.1, gamma_xz = 0.15, gamma_mz = 0.12

# Correlation between latent variables:
## cor_xm = 0.2, cor_xz = 0.3, cor_zm = 0.4

# Data simulation: 
## x, z, m, ey
cov_xmz_ey <- matrix(c(1, 0.1, 0.15, 0,
                       0.1, 1, 0.12, 0,
                       0.15, 0.12, 1, 0,
                       0, 0, 0, 0.481351), nrow = 4)
eta <- as.data.frame(MASS::mvrnorm(num_obs, mu = rep(0, 4), Sigma = cov_xmz_ey,
                     empirical = FALSE))
names(eta) <- c("x", "m", "z", "ey")

# xm, xz, mz
eta <- eta %>%
  mutate(xm = x*m,
         xz = x*z,
         mz = m*z)

# y
etay <- 0.3*eta$x + 0.4*eta$m + 0.2*eta$z + 0.1*eta$xm + 0.15*eta$xz + 0.12*eta$mz + eta$ey

# Observed Indicators
lambda_x <- c(0.9, 0.8, 0.7)
lambda_m <- c(0.85, 0.75, 0.65)
lambda_z <- c(0.8, 0.7, 0.6)
lambda_y <- c(0.75, 0.7, 0.65)

x_obs <- eta$x %*% t(lambda_x) + rnorm(num_obs * length(lambda_x))
m_obs <- eta$m %*% t(lambda_m) + rnorm(num_obs * length(lambda_m))
z_obs <- eta$z %*% t(lambda_z) + rnorm(num_obs * length(lambda_z))
y_obs <- etay %*% t(lambda_y) + rnorm(num_obs * length(lambda_z))

# Dataset: raw score
df <- cbind(x_obs, m_obs, z_obs, y_obs)
df <- as.data.frame(df)
names(df) <- c(paste0("x", 1:3), paste0("m", 1:3), paste0("z", 1:3), paste0("y", 1:3))

# Dataset: factor score using `get_fs()`
fs_dat <- get_fs(df, model = "x =~ x1 + x2 + x3
                              m =~ m1 + m2 + m3
                              z =~ z1 + z2 + z3
                              y =~ y1 + y2 + y3")

head(fs_dat)
```

```{r Demonstrating int_ind() function}
# With specified model
ind_1 <- get_fs_int(fs_dat,
                    fs_name = c("fs_x", "fs_m", "fs_z"),
                    se = c("fs_x_se", "fs_m_se", "fs_z_se"),
                    loading = c("x_by_fs_x", "m_by_fs_m", "z_by_fs_z"),
                    model = "fs_x:fs_m + fs_x:fs_z")
head(ind_1)

# Without specified model
ind_2 <- get_fs_int(fs_dat,
                    fs_name = c("fs_x", "fs_m", "fs_z"),
                    se = c("fs_x_se", "fs_m_se", "fs_z_se"),
                    loading = c("x_by_fs_x", "m_by_fs_m", "z_by_fs_z"))
head(ind_2)
```




