---
title: "tspa-interaction"
author: "Jimmy Zhang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tspa-interaction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r load packages}
library(lavaan)
library(semTools)
library(dplyr)
library(mnormt)
library(R2spa)
# library(nlsem)
```

# Introduction

Recent widely used approaches of estimating latent interaction effects are Unconstrained Product Indicator (UPI) and Reliability-Adjusted Product Indicator (RAPI). We proposed a new approach based on the Two-Stage Path Analysis (2S-PA) framework, 2S-PA-Interaction, and compared to the UPI and RAPI methods. 

We begin by simulating a dataset under a hypothetical structural model with two latent predictors, $\xi_{X}$ and $\xi_{M}$, and one observed dependent variable, $Y$. Each latent predicted is indicated by three items (i.e., [$x1$, $x2$, $x3$]; [$m1$, $m2$, $m3$]). The latent interaction term is demonstrated by a multiplicative term, $\xi_{XM}$. 

\begin{equation}
y = \gamma_{1}\xi_{X} + \gamma_{2}\xi_{M} + \gamma_{3}\xi_{XM} + \zeta,
\end{equation}

The $\zeta$ is the disturbance of regression model with a distribution of $\zeta \sim N(0, \Psi)$.

```{r data-generating function}
# Define a function of generating interaction data
generate_sem_data <- function(N, model, Alpha, Phi, Lambda, Gamma, Theta, SD_y) {
  # Generate scores for observed items: x1 - x3, m1 - m3
  eta_scores <- rmnorm(N, mean = Alpha, varcov = Phi) # Factor scores
  # Generate eta scores with interaction
  eta_scores_int <- cbind(eta_scores, eta_scores[,1]*eta_scores[,2])
  delta <- rmnorm(N, mean = rep(0, length(diag(Theta))), varcov = Theta) # Errors/Residuals of Indicators
  item_scores <- tcrossprod(eta_scores, Lambda) + delta # Item/Indicator scores

  y_scores <- tcrossprod(eta_scores_int, Gamma) + rnorm(N, 0, SD_y) # Y scores / DV scores

  # Parsing formula
  indicator_terms <- unlist(strsplit(gsub(" ", "",
                                          unlist(strsplit(model, "\n"))[grep("=~",
                                                                             unlist(strsplit(model, "\n")))]),
                                     split = "=~"))
  indicator_vars <- indicator_terms[grepl("+", indicator_terms, fixed = TRUE) == "FALSE"]
  indicator_num <- as.vector(unlist(lapply(strsplit(indicator_terms[grepl("+", indicator_terms, fixed = TRUE) == "TRUE"],
                                                    split = "+", fixed = TRUE), length)))

  df <- as.data.frame(cbind(item_scores, y_scores))
  names <- list()
  for (n in seq(length(indicator_vars))) {
    names[[n]] <- tolower(paste0(indicator_vars[n], seq(indicator_num[n])))
  }
  colnames(df) <- c(unlist(names), "Y")
  return(df)
}
```

```{r simulate data}
set.seed(4073)
# Define model parameters 
N <- 500
model <- "# Measurement Model
            X =~ x1 + x2 + x3
            M =~ m1 + m2 + m3
          # Structural Model
            Y ~ b1*X + b2*M + b3*X:M"
Alpha <- c(0,0) # Latent means of 0
Phi <- matrix(c(1, 0.6,
                0.6, 1), 
              nrow = 2) # Latent correlation of 0.6
Lambda <- cbind(c(1, 0.9, 0.75, rep(0, 3)),
                c(rep(0, 3), 1, 0.9, 0.75)) # Factor loadings
Gamma <- t(c(0.3, 0.3, 0.3)) # Path coefficients
Theta <- diag(c(1.32, 0.99, 0.69), nrow = 6) # Residual variance with reliability of 0.7
SD_y <- 0.7678542 # Error variance at cor_xm = 0.6

# Simulate Data
df <- generate_sem_data(N, model, Alpha, Phi, Lambda, Gamma, Theta, SD_y)
```

## UPI

The procedure of fitting the UPI model is adapted from Schoemann $\&$ Jorgensen (2021):

```{r manual upi}
# Step 1: Create product indicators
x_value <- paste0("x", 1:3)
m_value <- paste0("m", 1:3)
intNames <- paste0(rep(x_value, each = length(m_value)), m_value) # Create product indicator names
mod_df <- indProd(df, 
                  var1 = c("x1", "x2", "x3"),
                  var2 = c("m1", "m2", "m3"),
                  match = FALSE, 
                  meanC = T, 
                  residualC = F, 
                  doubleMC = T,
                  namesProd = intNames) # using the DMC strategy

# Step 2: Draft the model syntax
mod_upi <- '# Measurement Model
            X =~ x1 + x2 + x3
            M =~ m1 + m2 + m3
            XM =~ x1m1 + x1m2 + x1m3 + 
                  x2m1 + x2m2 + x2m3 +
                  x3m1 + x3m2 + x3m3
                   
            # Structural Model
            Y ~ b1*X + b2*M + b3*XM
            
            # Covariances between the same indicators are constrained to equality
              x1m1 ~~ th1*x1m2 + th1*x1m3 
              x1m2 ~~ th1*x1m3 
              x2m1 ~~ th2*x2m2 + th2*x2m3 
              x2m2 ~~ th2*x2m3 
              x3m1 ~~ th3*x3m2 + th3*x3m3 
              x3m2 ~~ th3*x3m3
              x1m1 ~~ th4*x2m1 + th4*x3m1 
              x2m1 ~~ th4*x3m1 
              x1m2 ~~ th5*x2m2 + th5*x3m2 
              x2m2 ~~ th5*x3m2 
              x1m3 ~~ th6*x2m3 + th6*x3m3 
              x2m3 ~~ th6*x3m3

            # Define Standardized Coefficients
              X ~~ v1*X
              M ~~ v2*M
              beta1 := b1*sqrt(v1)
              beta2 := b2*sqrt(v2)
              beta3 := b3*sqrt(v1)*sqrt(v2)'

# Step 3: Fit the model
fit_upi <- sem(mod_upi, mod_df)
summary(fit_upi, fit.measure = TRUE, standardized = TRUE)
```

The steps above have been wrapped up into a function `upi`. Here is an example demonstration:

```{r function upi}
# Draft the model syntax
mod_upi <- '# Measurement Model
              X =~ x1 + x2 + x3
              M =~ m1 + m2 + m3
            # Structural Model
              Y ~ b1 * X + b2 * M + b3 * X:M
            # Define Standardized Coefficients
              X ~~ v1*X
              M ~~ v2*M
              
              beta1 := b1*sqrt(v1)
              beta2 := b2*sqrt(v2)
              beta3 := b3*sqrt(v1)*sqrt(v2)'

# Fit the model using upi()
fit_upi <- upi(mod_upi, df, mode = "all") # All-pair UPI
summary(fit_upi, fit.measure = TRUE, standardized = TRUE)

# Matched-pair
fit_upi_matched <- upi(mod_upi, df, mode = "match")
summary(fit_upi_matched, fit.measure = TRUE, standardized = TRUE)
```

## RAPI

The steps of applying RAPI method is based on technical details in Hsiao et al. (2018).

```{r manually fit rapi loading, warning=FALSE}
# Step 1: Prepare mean-centered df
df_centered <- df %>%
  mutate(across(c(x1:m3), ~ . - mean(.)))

# Step 2: Using sum scores as the single indicators
df_centered$x_mean <- rowSums(df_centered[, c("x1", "x2", "x3")])/3
df_centered$m_mean <- rowSums(df_centered[, c("m1", "m2", "m3")])/3

# Step 3: Create the interaction indicator
df_centered$xm <- df_centered$x_mean*df_centered$m_mean

# Step 4: Calculate reliability of indicators
X_rel <- psych::alpha(df_centered[, c("x1", "x2", "x3")])$total$raw_alpha
M_rel <- psych::alpha(df_centered[, c("m1", "m2", "m3")])$total$raw_alpha

# Step 5: Draft the model syntax
mod_rapi <- 'X =~ 1 * x_mean
             M =~ 1 * m_mean
             XM =~ 1 * xm
             
             # Error variance
             x_mean ~~ ev_x * x_mean
             m_mean ~~ ev_m * m_mean
             xm ~~ ev_xm * xm
             
             # Regression
             Y ~ b1 * X + b2 * M + b3 * XM
             
             # Latent variance
             X ~~ v_x * X
             M ~~ v_m * M
             XM ~~ v_xm * XM
             
             # Error Constraints
             ev_x == (1 - X_rel) * v_x / X_rel
             ev_m == (1 - M_rel) * v_m / M_rel
             ev_xm == ev_x * v_m + ev_m * v_x + ev_x * ev_m
             
             # x-standardized interaction
             beta1 := b1 * sqrt(v_x)
             beta2 := b2 * sqrt(v_m)
             beta3 := b3 * sqrt(v_x) * sqrt(v_m)'

# Substitute in the actual reliability estimates
mod_rapi <- gsub("X_rel", replacement = X_rel, x = mod_rapi)
mod_rapi <- gsub("M_rel", replacement = M_rel, x = mod_rapi)

# Step 6: Fit the RAPI model
rapi_fit <- sem(mod_rapi, data = df_centered)
summary(rapi_fit)
```

```{r manually fit rapi lat_var, warning=FALSE}
# Step 1: Retrieve factor loadings from fitted upi
mod <- 'X =~ x1 + x2 + x3
        M =~ m1 + m2 + m3'
PE <- parameterEstimates(sem(mod, df))

# Step 2: Caculate sum of factor loadings, error variance
# Sum of factor loadings
X_lambda <- sum(PE[match(c("x1", "x2", "x3"), PE$rhs), "est"])
M_lambda <- sum(PE[match(c("m1", "m2", "m3"), PE$rhs), "est"])
  
# Sum of error variances
PE_vcov <- PE[PE$op == "~~", ]
x_ev <- sum(PE_vcov[match(c("x1", "x2", "x3"), PE_vcov$rhs), "est"])
m_ev <- sum(PE_vcov[match(c("m1", "m2", "m3"), PE_vcov$rhs), "est"])
  
# Mean of item scores
df_centered <- df %>%
  mutate(across(c(x1:m3), ~ . - mean(.)))
df_centered$x_sum <- rowSums(df_centered[, c("x1", "x2", "x3")])
df_centered$m_sum <- rowSums(df_centered[, c("m1", "m2", "m3")])
  
# Step 3: Create the interaction indicator
df_centered$xm <- df_centered$x_sum*df_centered$m_sum

# Step 4: Calculate reliability of indicators and error constraint for the interaction term
x_rel <- psych::alpha(df_centered[, c("x1", "x2", "x3")])$total$raw_alpha
m_rel <- psych::alpha(df_centered[, c("m1", "m2", "m3")])$total$raw_alpha
int_ev <- x_rel*var(df_centered$x_sum, na.rm = T)*(1 - m_rel)*var(df_centered$m_sum, na.rm = T) +
    m_rel*var(df_centered$m_sum, na.rm = T)*(1 - x_rel)*var(df_centered$x_sum, na.rm = T) +
    (1 - m_rel)*var(df_centered$m_sum, na.rm = T)*(1 - x_rel)*var(df_centered$x_sum, na.rm = T)

# Step 5: Draft the model syntax
mod_rapi <- 'X =~ X_lambda * x_sum
             M =~ M_lambda * m_sum
             XM =~ X_lambda * M_lambda * xm
             
             # Error variance
             x_sum ~~ x_ev * x_sum
             m_sum ~~ m_ev * m_sum
             xm ~~ int_ev * xm
             
             # Regression
             Y ~ X + M + b3 * XM
             
             # Latent variance
             X ~~ v_x * X
             M ~~ v_m * M
             XM ~~ v_xm * XM
             
             # x-standardized interaction
             beta3 := b3 * sqrt(v_x) * sqrt(v_m)'

# Substitute in the actual reliability estimates
# Define patterns and their replacements
replacements <- list(
  X_lambda = X_lambda,
  M_lambda = M_lambda,
  x_ev = x_ev,
  m_ev = m_ev,
  int_ev = int_ev
)

# Iterate over the replacements and update the model string
for (pattern in names(replacements)) {
  mod_rapi <- gsub(pattern, replacements[[pattern]], mod_rapi)
}

# Step 6: Fit the RAPI model
rapi_fit <- sem(mod_rapi, data = df_centered)
summary(rapi_fit)
```

The RAPI method using factor loading constraints as 1s above have been wrapped up into a function `rapi`. Here is an example demonstration:

```{r function rapi}
# Draft the model syntax
mod_rapi <- '# Measurement Model
              X =~ x1 + x2 + x3
              M =~ m1 + m2 + m3
            # Structural Model
              Y ~ b1*X + b2*M + b3*X:M
            # Define Standardized Coefficients
              X ~~ v1*X
              M ~~ v2*M
              beta1 := b1*sqrt(v1)
              beta2 := b2*sqrt(v2)
              beta3 := b3*sqrt(v1)*sqrt(v2)'

# Fit the model using rapi()
fit_rapi <- rapi(model = mod_rapi, data = df) 
summary(fit_rapi, fit.measure = TRUE, standardized = TRUE)
```

# 2S-PA

The 2S-PA interaction model is based on the 2S-PA framework proposed by Lai & Hsiao (2022).

```{r manually fit 2s-pa model}
# Step 1: Get factor scores
fs_dat <- get_fs(df,
                 model ='X =~ x1 + x2 + x3
                         M =~ m1 + m2 + m3',
                 method = "Bartlett",
                 std.lv = TRUE)

# Step 2: Obtain mean-centered factor product
fs_dat$fs_int <- fs_dat$fs_X * fs_dat$fs_M
fs_dat$fs_int <- fs_dat$fs_int - mean(fs_dat$fs_int)
Y <- df$Y
fs_dat <- cbind(fs_dat, Y)

# Step 3: Obtain the standard error of the interaction term
fs_dat$fs_int_se <- sqrt(1 * fs_dat$fs_X_se^2 + 1 * fs_dat$fs_M_se^2 + fs_dat$fs_X_se^2*fs_dat$fs_M_se^2)

# Step 4: 
mod_tspa = '# latent variables (indicated by factor scores)
            X =~ fs_X
            M =~ fs_M
            XM =~ fs_int
            
            # constrain the errors
            fs_X ~~ rel_X * fs_X
            fs_M ~~ rel_M * fs_M
            fs_int ~~ rel_int * fs_int
            
            # regressions
            Y ~ b1*X + b2*M + b3*XM

            X ~~ v1*X
            M ~~ v2*M
            XM ~~ v3*XM
              
            # Define Standardized Coefficients
            beta1 := b1 * sqrt(v1)
            beta2 := b2 * sqrt(v2)
            beta3 := b3 * sqrt(v1) * sqrt(v2)'

mod_tspa <- gsub("rel_X", replacement = fs_dat$fs_X_se[1]^2, x = mod_tspa)
mod_tspa <- gsub("rel_M", replacement = fs_dat$fs_M_se[1]^2, x = mod_tspa)
mod_tspa <- gsub("rel_int", replacement = fs_dat$fs_int_se[1]^2, x = mod_tspa)

fit_tspa <- sem(mod_tspa, data = fs_dat)
summary(fit_tspa)
```

```{r function tspa}
fit_tspa <- tspa(model = "Y ~ b1*X + b2*M + b3*X:M
                          beta1 := b1 * sqrt(v1)
                          beta2 := b2 * sqrt(v2)
                          beta3 := b3 * sqrt(v1) * sqrt(v2)",
                   data = fs_dat,
                   se = list(X = fs_dat$fs_X_se[1],
                             M = fs_dat$fs_M_se[1]))
summary(fit_tspa)
```






